Ship-based lidar evaluation of Southern Ocean clouds in the storm-resolving general circulation model ICON, and the ERA5 and MERRA-2 reanalyses

### 2024-12-13 manuscript_agu.tex
Update on Overleaf.

\subsection{CERES}

@@

In all situations, even when the models overestimate cloud occurrence at some altitudes, they always substantially underestimate the total cloud fraction. ICON can be generally characterized as substantially overestimating cloud occurrence below 1 km and underestimating above, underestimating the total cloud fraction, and showing the greatest biases in conditions of weak stability and non-cyclonic conditions. ICON also has a peak cloud occurrence at higher altitudes than observations (500~m vs. near the surface), and correspondingly, its LCL tends to be higher. MERRA-2 can be generally characterized as underestimating cloud occurrence at nearly all altitudes as well as the total cloud fraction, but mostly above and below 500~m (the peak at 500~m is well represented). MERRA-2 displays the largest errors relative to observations in the low-latitude SO zone and in situations of weak stability. ERA5 can be generally characterized as representing cloud occurrence correctly above about 1--1.5~km, overestimating between 500m and 1km, but underestimating near-surface cloud occurrence (0--500~m). The total cloud fraction is strongly underestimated in all situations. ERA5 has a tendency towards underestimation in the low-latitude SO zone and situations of weak stability; conversely, it overestimates in the high-latitude SO zone and conditions of strong stability.

@@

In Fig.~\ref{fig:cloud-occurrence}, we also display the mean outgoing shortwave and longwave top of atmosphere radiation. In observations, these come from daily mean CERES measurements averaged over the voyage tracks or a station location, whereas in the models they come from daily means of TOA radiation in the model output averaged over the same location and time periods. In the free-running ICON model, the time period is mapped onto the available years, as explained in Section~\ref{sec:icon}.

In the general case (Fig.~\ref{fig:cloud-occurrence}a), ICON underestimates the outgoing SW radiation by 26 Wm$^{-2}$ and the MERRA-2 and ERA5 reanalyses overestimate it by 6 and 14 Wm$^{-2}$, respectively. While in ICON, this is in line with the underestimated total cloud fraction of 10\%, in the reanalyses this is the opposite result to that expected from the underestimated total cloud fraction of about 20\%. The likely explanation is an overestimated cloud albedo, compensating for the lack of cloud area. We note that the the radiative transfer calculations used in the lidar simulator mean that the impact of both cloud phase and cloud fraction are convolved which may help explain this result.

@@

Our main finding corroborates previous findings of large boundary layer cloud biases in models and their subsequent effect on the radiative transfer. For example, low- and mid-level clouds in the cold-air sector of cyclones were identified as being responsible for most of the SW bias in \cite{Bodas-Salcedoetal2012}. This understanding was refined in \cite{Bodas-Salcedoetal2014} which highlighted that the SW bias was associated with an incorrectly simulated mid-level cloud regime, which occurred in regions where cloud with tops at mid-level and low-levels occurred. Our results align less well with more recent work by \cite{ramadoss2024} which shows persistent shortwave radiative bias over the Southern Ocean are associated with incorrect cloud phase representation. While \cite{fiddes2024} suggest biases in the liquid water path are the largest contributor to the cloud radiative bias over the Southern Ocean. Our general finding applies to the new GSRM ICON, but the biases are generally lower than in the reanalyses, despite the reanalyses having the advantage of assimilation of the observed meteorological conditions. The GSRM has, on the other hand, the advantage of a much higher spatial resolution and, to a limited extent, explicit calculation of traditionally subgrid-scale processes such as convection.

@@

The radiosonde observations indicate that the LCL is too high in ICON and reanalyses, which is probably responsible for the higher peak of clouds in the models and the lack of near-surface clouds or fog. The radiosonde comparison, however, does not seem to explain cloud biases at higher altitudes which is perhaps suggestive of biases in the influence of the liquid water path in teh models relative to the reality observed by the models. MERRA-2 is too moist at all heights. ICON also exhibits smaller internal variability than the radiosonde observations. Overall, the radiosonde comparison only partially explains the identified cloud biases, and other physical causes are likely contributing. This warrants further investigation, especially of ocean--atmosphere fluxes, shallow convection, and boundary layer turbulence. The lack of parameterized subgrid-scale convection in ICON could be a substantial issue even at the 5-km resolution.

>>>>

\subsection{CERES}
\label{sec:ceres}

@@

In all situations, even when the models overestimate cloud occurrence at some altitudes, they always substantially underestimate the total cloud fraction. ICON can be generally characterized as substantially overestimating cloud occurrence below 1 km and underestimating above, underestimating the total cloud fraction, and showing the greatest biases in conditions of weak stability and non-cyclonic conditions. ICON also has a peak cloud occurrence at higher altitudes than observations (500~m vs. near the surface), and correspondingly, its LCL tends to be higher. MERRA-2 can be generally characterized as underestimating cloud occurrence at nearly all altitudes as well as the total cloud fraction, but mostly above and below 500~m (the peak at 500~m is well represented). MERRA-2 displays the largest errors relative to observations in the low-latitude SO zone and in situations of weak stability. ERA5 can be generally characterized as representing cloud occurrence correctly above about 1.5~km, overestimating between 500~m and 1~km, but underestimating near-surface cloud occurrence (0--500~m). The total cloud fraction is strongly underestimated in all situations. ERA5 has a tendency towards underestimation in the low-latitude SO zone and situations of weak stability; conversely, it overestimates in the high-latitude SO zone and conditions of strong stability.

@@

In Fig.~\ref{fig:cloud-occurrence}, we also display the mean outgoing shortwave and longwave top of atmosphere radiation, whose calculation is described in Section~\ref{sec:ceres}. In observations, these come from daily mean CERES measurements averaged over the voyage tracks or a station location, whereas in the models they come from daily means of TOA radiation in the model output averaged over the same location and time periods. In the free-running ICON model, the time period is mapped onto the available years, as explained in Section~\ref{sec:icon}.

In the general case (Fig.~\ref{fig:cloud-occurrence}a), ICON underestimates the outgoing SW radiation by 26 Wm$^{-2}$ and the MERRA-2 and ERA5 reanalyses overestimate it by 6 and 14 Wm$^{-2}$, respectively. While in ICON, this is in line with the underestimated total cloud fraction of 10\%, in the reanalyses this is the opposite result to that expected from the underestimated total cloud fraction of about 20\%. The likely explanation is an overestimated cloud albedo, compensating for the lack of cloud area.

We note that the the radiative transfer calculations used in the lidar simulator mean that the impact of both cloud phase and cloud fraction are convolved to produce the cloud mask. Therefore, the cloud occurrence is not affected by any cloud phase biases as long as the cloud is optically thick enough to be detected and the laser signal is not too attenuated. However, a combination of underestimated total cloud fraction and overestimated outgoing SW at TOA is indicative of an overestimated cloud albedo due to either cloud liquid and ice water content, cloud phase, droplet or ice crystal size distribution, shape or orientation of ice crystals, or cloud overlap.

@@

Our main finding corroborates previous findings of large boundary layer cloud biases in models and their subsequent effect on the radiative transfer. For example, low- and mid-level clouds in the cold-air sector of cyclones were identified as being responsible for most of the SW bias in \cite{bodas-salcedo2012}. This understanding was refined in \cite{bodas-salcedo2014}, which highlighted that the SW bias was associated with an incorrectly simulated mid-level cloud regime, which occurred in regions where cloud with tops at mid-level and low-levels occurred. Our results align less well with more recent work by \cite{ramadoss2024} which shows persistent shortwave radiative bias over the Southern Ocean are associated with incorrect cloud phase representation. While \cite{fiddes2024} suggest biases in the liquid water path are the largest contributor to the cloud radiative bias over the Southern Ocean. Our general finding applies to the new GSRM ICON, but the biases are generally lower than in the reanalyses, despite the reanalyses having the advantage of assimilation of the observed meteorological conditions. The GSRM has, on the other hand, the advantage of a much higher spatial resolution and, to a limited extent, explicit calculation of traditionally subgrid-scale processes such as convection.

@@

The radiosonde observations indicate that the LCL is too high in ICON and reanalyses, which is probably responsible for the higher peak of clouds in the models and the lack of near-surface clouds or fog. The radiosonde comparison, however, does not seem to explain cloud biases at higher altitudes, which is perhaps suggestive of biases in the influence of the liquid water path in the models relative to the reality. MERRA-2 is too moist at all heights. ICON also exhibits smaller internal variability than the radiosonde observations. Overall, the radiosonde comparison only partially explains the identified cloud biases, and other physical causes are likely contributing. This warrants further investigation, especially of ocean--atmosphere fluxes, shallow convection, and boundary layer turbulence. The lack of parameterized subgrid-scale convection in ICON could be a substantial issue even at the 5-km resolution.
